Modelling last glacial cycle ice dynamics in the Alps

### 2017-11-05 articles/cycle/alps-cycle.tex
Turn large maps sideways.
Make a few more previously ignore fixes.

    From the mid-twentieth
    century, palaeo-climate records extracted from deep sea sediments and
    ice cores have provided a much more detailed picture of the Earth
    environmental history \citep[e.g.,][]{Emiliani.1955,
    Shackleton.Opdyke.1973, Dansgaard.etal.1993, Augustin.etal.2004},
    indicating tens of glacial and
    interglacial periods. During the last 800\,000 years (800\,ka), these
    \emph{glacial cycles}
    have succeeded each other with a periodicity about 100\,ka \citep{Hays.etal.1976,
    Augustin.etal.2004}. Nevertheless, this global signal is largely governed
    by the North American and Eurasian ice sheet complexes. It is thus unclear
    whether glacier advances and retreats in the Alps were in pace with global
    sea-level fluctuations.

@@

    The palaeo-climate forcing was adapted (Sect.~\ref{sec:paltemp})
    to model this maximum configuration in broad agreement with the geological
    reconstruction of the LGM extent (Fig.~\ref{fig:lgmvel}a, solid red line),
    yet local discrepancies remain. As already outlined for low resolution
    runs (Sect.~\ref{sec:footprints}), the extent of glaciation in the
    north-western Alps, including the Rhone Glacier complex, the Jura ice cap,
    and the Lyon Lobe, is underestimated in the model results
    (Fig.~\ref{fig:lgmvel}a). On the other hand, the model yields excessive ice
    cover in the eastern Alps, where the Mur, Drava and Sava glaciers are
    modelled to extend tens of km beyond the mapped ice limits
    (Fig.~\ref{fig:lgmvel}a). As previously discussed, these discrepancies
    might indicate that the
    LGM climate was characterized by an east-west gradient in temperature
    \citep[cf.][]{Heyman.etal.2013} or precipitation
    \citep[cf.][]{Wu.etal.2007} anomalies relative to present, or an east-west
    gradient in variables that are not accounted for by the PDD model.

@@

    Our results indicate that even for a relatively small ice sheet like that
    found in the Alps, the LGM glacier extent corresponds to a transient
    stage, while ka-scale inherent differences in its timing can
    result from complex glacier dynamics. Further differences in LGM timing
    caused by heterogeneous climatic anomalies, such as precipitation
    anomaly gradients, not included in our model set-up, may have
    counterbalanced or enhanced those modelled here.

@@

    \begin{figure*}
      \centerline{\includegraphics{alpcyc_hr_lgmvel}}
      \caption{%
        \textbf{(a)} Modelled bedrock topography (grey), ice surface topography
        (200\,m contours), and ice surface velocity (blue) in the Alps
        24.57\,ka before present, corresponding to the maximum modelled ice
        cover. Modelled self-sustained ice domes (triangles) and some major
        transfluences (crosses) are shown. Modelled Last Glacial Maximum
        (LGM) ice extent (dashed orange line) and geomorphological
        reconstruction \citep[solid red line,][]{Ehlers.etal.2011}. The
        background map consists of depressed SRTM \citep{Jarvis.etal.2008}
        topography and Natural Earth Data \citep{Patterson.Kelso.2017}.
        \textbf{(b)} Temperature offset time-series from the EPICA ice core
        used as palaeo-climate forcing for the ice flow model \citep[black
        curve,][]{Jouzel.etal.2007}, and modelled total ice volume through the
        last glacial cycle (120--0\,ka), expressed in meters of sea level
        equivalent (m~s.l.e., blue curve). Gray fields indicate Marine
        Oxygen Isotope Stage (MIS) boundaries for MIS~2 and MIS~4 according to
        a~global compilation of benthic \chem{\delta^{18}O} records
        \citep{Lisiecki.Raymo.2005}.
        \todo{Rename Ivrea to Dora Baltea.}}
      \label{fig:lgmvel}
    \end{figure*}

    \begin{figure*}
      \centerline{\includegraphics{alpcyc_hr_timing}}
      \caption{%
        \textbf{(a)} Timing of the LGM given by the modelled age of maximum ice
        thickness throughout the entire simulation (colour mapping) and
        corresponding, ice surface elevation (200\,m contours).
        \textbf{(b)} Temperature offset time-series from the EPICA ice core
        used as palaeo-climate forcing for the ice flow model (black curve),
        and modelled glacierized area during the LGM (coloured curve). The LGM
        is here modelled as a time-transgressive event.
        \todo{Rename Ivrea to Dora Baltea.}}
      \label{fig:timing}
    \end{figure*}

>>>>

    From the mid-twentieth
    century, palaeo-climate records extracted from deep sea sediments and
    ice cores have provided a much more detailed picture of the Earth
    environmental history \citep[e.g.,][]{Emiliani.1955,
    Shackleton.Opdyke.1973, Dansgaard.etal.1993, Augustin.etal.2004},
    indicating tens of glacial and interglacial periods. During the last
    800\,000 years before present (800\,ka), these \emph{glacial cycles}
    have succeeded each other with a periodicity about 100\,ka \citep{Hays.etal.1976,
    Augustin.etal.2004}. Nevertheless, this global signal is largely governed
    by the North American and Eurasian ice sheet complexes. It is thus unclear
    whether glacier advances and retreats in the Alps were in pace with global
    sea-level fluctuations.

@@

    The palaeo-climate forcing was adapted (Sect.~\ref{sec:paltemp})
    to model the maximum configuration of the Rhine Glacier in broad agreement
    with the geological
    reconstruction of the LGM extent (Fig.~\ref{fig:lgmvel}a, solid red line),
    yet discrepancies remain elsewhere. As already outlined for low resolution
    runs (Sect.~\ref{sec:footprints}), the extent of glaciation in the
    north-western Alps, including the Rhone Glacier complex, the Jura ice cap,
    and the Lyon Lobe, is underestimated in the model results
    (Fig.~\ref{fig:lgmvel}a). On the other hand, the model yields excessive ice
    cover in the eastern Alps, where the Mur, Drava and Sava glaciers are
    modelled to extend tens of km beyond the mapped ice limits
    (Fig.~\ref{fig:lgmvel}a). As previously discussed, these discrepancies
    might indicate that the
    LGM climate was characterized by an east-west gradient in temperature
    \citep[cf.][]{Heyman.etal.2013} or precipitation
    \citep[cf.][]{Wu.etal.2007} anomalies relative to present, or an east-west
    gradient in variables that are not accounted for by the PDD model.

@@

    Our results indicate that even for a relatively small ice sheet like that
    found in the Alps, the LGM glacier extent corresponds to a transient stage,
    while millennial scale, spatial differences in its timing can result not
    only from climate variability but also from complex glacier dynamics.
    Heterogeneous climatic anomalies, such as temperature or precipitation
    anomaly gradients, not included in our model set-up, have perhaps induced
    further spatial differences in the timing of the LGM Alpine ice sheet,
    which may have counterbalanced or enhanced those modelled here.

@@

    \begin{sidewaysfigure*}
      \centerline{\includegraphics{alpcyc_hr_lgmvel}}
      \caption{%
        \textbf{(a)} Modelled bedrock topography (grey), ice surface topography
        (200\,m contours), and ice surface velocity (blue) in the Alps
        24.57\,ka before present, corresponding to the maximum modelled ice
        cover. Modelled self-sustained ice domes (triangles) and some major
        transfluences (crosses) are shown. Modelled Last Glacial Maximum
        (LGM) ice extent (dashed orange line) and geomorphological
        reconstruction \citep[solid red line,][]{Ehlers.etal.2011}. The
        background map consists of depressed SRTM \citep{Jarvis.etal.2008}
        topography and Natural Earth Data \citep{Patterson.Kelso.2017}.
        \textbf{(b)} Temperature offset time-series from the EPICA ice core
        used as palaeo-climate forcing for the ice flow model \citep[black
        curve,][]{Jouzel.etal.2007}, and modelled total ice volume through the
        last glacial cycle (120--0\,ka), expressed in meters of sea level
        equivalent (m~s.l.e., blue curve). Gray fields indicate Marine
        Oxygen Isotope Stage (MIS) boundaries for MIS~2 and MIS~4 according to
        a~global compilation of benthic \chem{\delta^{18}O} records
        \citep{Lisiecki.Raymo.2005}.
        \todo{Enlarge to full width, rename Ivrea to Dora Baltea.}}
      \label{fig:lgmvel}
    \end{sidewaysfigure*}

    \begin{sidewaysfigure*}
      \centerline{\includegraphics{alpcyc_hr_timing}}
      \caption{%
        \textbf{(a)} Timing of the LGM given by the modelled age of maximum ice
        thickness throughout the entire simulation (colour mapping) and
        corresponding, ice surface elevation (200\,m contours).
        \textbf{(b)} Temperature offset time-series from the EPICA ice core
        used as palaeo-climate forcing for the ice flow model (black curve),
        and modelled glacierized area during the LGM (coloured curve). The LGM
        is here modelled as a time-transgressive event.
        \todo{Enlarge to full width, rename Ivrea to Dora Baltea.}}
      \label{fig:timing}
    \end{sidewaysfigure*}
